# Setting up a `phylochemistry` analysis

`phylochemistry` is not just a set of functions that helps with analyses. It is also a conceptual and organizational framework that helps you perform analyses efficiently and with versatility. It takes a few minutes to understand and set up a phylochemistry analysis, but that investment will be paid back many-fold over the course of a project's lifetime.

## Conventions in `phylochemsitry`

### File paths and directory paths

`phylochemistry` is made to run both on data from single sources (for example, one species) or multiple sources (for example, multiple species) at the same time. This means that sometimes a function will run using a single file, or sometimes you will want to feed it multiple files. In cases where a function wants a single file, it will request a "file_path", which is an explicit path to a single file (for example, "Desktop/project/file.csv"). Other times, a function will want a directory instead of a specific file, in which case it will request a "directory_path". This is a path to a directory and it ends with a slash (for example, "Desktop/project/").

### Monolists

One way to think of `phylochemistry` is as a set of tools that enables you to integrate and analyze raw data stored on your hard drive. Let's use the example dataset included with `phylochemistry` to explore this idea. First, let's load the phylochemistry library and set up our phylochemistry directory:

```{r}
library(phylochemistry)
phylochemistry_example_dir <- "/Users/lucasbusta/Desktop/phylochemistry_example_data/"
phylochemistry_example_dir
```

Now, let's see what's in that directory
```{r}
dir(phylochemistry_example_dir)
```

We can see that there's a folder called monolists. What's in there?
```{r}
dir(paste0(phylochemistry_example_dir, "monolists/"))
```

### Storing your data as a list of lists

There are lots of different kinds of data structures. `phylochemistry` was designed with a very simple and flexible data structure in mind: a heirarchy of lists. These are very easy to create and modify:

```{r}
my_analysis <- list()
my_analysis$chromatography_samples <- readMonolist(monolist_in_path = paste0(phylochemistry_example_dir, "monolists/chemistry_samples.csv"))
my_analysis$chromatography_samples
```

This is a list of chromatography samples and how much of two different types of standards are in each.

### Nomenclature
